Viscous Erosion in Porous Media

### 2018-02-22 docs/methods/erosion_methods.tex
NM

For numerical stability, we modify the erosion law \eqref{eqn:erosionModel} to include a term that depends on the local curvature $\kappa$. The idea is to prevent regions of very high
curvature from developing, as these regions would reduce the accuracy of our fluid solver. We thus modify the normal velocity of each interface as follows
\begin{align}
  \label{eqn:Vn}
  \Vn = \atau + \eps \mean{\atau} \left( \frac{L }{2\pi} \kappa - 1
  \right).
\end{align}
The second term on the right is the curvature-dependent modification, the strength of which is controlled by the small parameter $\eps \ll 1$. Here, $L$ is the total arc length of the body, and $\mean{\cdot}$ indicates a boundary mean. This term causes regions of high curvature to recede more rapidly, thus smoothing any high-frequency oscillations that my exist. Since a closed body has mean curvature $\mean{\kappa} = 2\pi / L$, this smoothing term has mean zero and and thus preserves area. Therefore, the only source of material loss is the shear-stress itself. In addition, we have scaled the smoothing term on the mean absolute shear, $\mean{\atau}$, so that the ratio of smoothing-to-erosion always remains fixed, even as bodies vanish.

@@

The main idea is to introduce an artificial tangential velocity that
keeps collocation points equally spaced with respect to arc length;
i.e.~the grid of $\alpha$ should always remain equally spaced. The
tangential velocity that does the trick satisfies~\cite{hou-low-she1994}
\begin{align*}
\pderiv{\Vs}{\alpha} = \thalpha \Vn - \mean{\thalpha \Vn}.
\end{align*}
The interface evolution is then given by
\begin{align*}
\dot{\xx}(t) = \Vn \nn + \Vs \ss.
\end{align*}
This evolution equation can be recast in the variables $\theta$ and $L$
as
\begin{align*}
\tderiv{L}{t} &= - 2 \pi \mean{\thalpha \Vn}, \\
\pderiv{\theta}{t} &= \frac{2 \pi}{L} \left( \pderiv{\Vn}{\alpha} +
\thalpha \Vs \right).
\end{align*}
Inserting~\eqref{eqn:Vn2} for the normal velocity gives
\begin{align}
\label{eqn:Levo1}
\tderiv{L}{t} &= - 2 \pi \mean{\thalpha \Vn}, \\
\label{eqn:thetaevo1}
\pderiv{\theta}{t} &= \frac{2\pi}{L} \left(
\eps \mean{\atau} \ppd{\theta}{\alpha} + \pderiv{\atau}{\alpha} +
\thalpha \Vs \right).
\end{align}
Note that stiffest term in~\eqref{eqn:thetaevo1} involving the
second-order derivative is {\em quasi-linear}. This is the key to the
{\thL} method, and allows us to use very stable, implicit methods for
this stiff term that are described in Section~\ref{sec:timeStepping}.

>>>>

For numerical stability, we modify the erosion law \eqref{eqn:erosionModel} to include a term that depends on the local curvature $\kappa$. The idea is to prevent regions of very high
curvature from developing, as these regions would compromise the accuracy and stability of our fluid solver. We thus modify the normal velocity of each interface as follows
\begin{align}
  \label{eqn:Vn}
  \Vn = \atau + \eps \mean{\atau} \left( \frac{L }{2\pi} \kappa - 1
  \right).
\end{align}
The second term on the right is the curvature-dependent modification, the strength of which is controlled by the small parameter $\eps \ll 1$. Here, $L$ is the total arc length of the body, and $\mean{\cdot}$ indicates a boundary mean. This term causes regions of high curvature to recede more rapidly, thus smoothing any high-frequency oscillations that my exist. Since a closed body has mean curvature $\mean{\kappa} = 2\pi / L$, this smoothing term has mean zero and and thus preserves area. Therefore, the only source of material loss is the shear-stress itself. In addition, we have scaled the smoothing term on the mean absolute shear, $\mean{\atau}$, so that the ratio of smoothing-to-erosion always remains fixed, even as bodies vanish.

@@

The main idea is to introduce an artificial tangential velocity that
keeps collocation points equally spaced with respect to arc length;
i.e.~the grid of $\alpha$ should always remain equally spaced. The
tangential velocity that does the trick satisfies~\cite{hou-low-she1994}
\begin{align*}
\pderiv{\Vs}{\alpha} = \thalpha \Vn - \mean{\thalpha \Vn}.
\end{align*}
The interface evolution is then given by
\begin{align*}
\dot{\xx}(t) = \Vn \nn + \Vs \ss.
\end{align*}
This evolution equation can be recast in the variables $\theta$ and $L$
as
\begin{align*}
\tderiv{L}{t} &= - 2 \pi \mean{\thalpha \Vn}, \\
\pderiv{\theta}{t} &= \frac{2 \pi}{L} \left( \pderiv{\Vn}{\alpha} +
\thalpha \Vs \right).
\end{align*}
Inserting~\eqref{eqn:Vn2} for the normal velocity gives
\begin{align}
\label{eqn:Levo1}
\tderiv{L}{t} &= - 2 \pi \mean{\thalpha \Vn}, \\
\label{eqn:thetaevo1}
\pderiv{\theta}{t} &= \frac{2\pi}{L} \left(
\eps \mean{\atau} \ppd{\theta}{\alpha} + \pderiv{\atau}{\alpha} +
\thalpha \Vs \right).
\end{align}
System \eqref{eqn:Levo1}--\eqref{eqn:thetaevo1} comprise the main evolution equations that we will solve numerically in Section~\ref{sec:timeStepping}.
